---
title: "ML models exercise"
---

First to install packages
```{r}
library(here)
library(dplyr)
library(ggplot2)
library(tidymodels)
library(plyr)
library(corrplot)
library(ranger)
library(glmnet)
```

Then to load data and set seed
```{r}
data_location <- here::here("ml-models-exercise", "clean.rds")
df <- readRDS(data_location)
rngseed <- 1234
set.seed(rngseed)
```

Change race 7 and 88 to category 3
```{r}
df$RACE <- revalue(df$RACE, c("7" = "3", "88" = "3"))
```

Make a pairwise correlation plot for continuous variables
```{r}
cor_df <- df %>% select(-DOSE, -SEX, -RACE)
cor_plot1 <- corrplot::corrplot(cor(cor_df, use = "complete.obs"), method = "number", type = "upper")
cor_plot1
```
Nothing seems super correlated!

Compute BMI
I am going to assume height is in meters, and weight in kg
Plugging random values into a calculator and converting into imperial gives me weights and heights that make sense
The formula I found is weight(kg)/height(m)^2
```{r}
df1 <- df %>% mutate(BMI = WT/(HT*HT))
```

Create 3 models using all variables
1. linear
2. Lasso
3. random forest
Make predictions for entire data set
Make observed vs. predicted plots
```{r}
lm_mod1 <- linear_reg()
lm_fit1 <- lm_mod1 %>% fit(Y ~ DOSE * AGE * SEX * WT * HT * RACE, data = df1)

ls_mod1 <- linear_reg(penalty = .1) %>% set_engine("glmnet")
ls_fit1 <- ls_mod1 %>% fit(Y ~ DOSE * AGE * SEX * WT * HT * RACE, data = df1)

rf_mod1 <- rand_forest(trees = 1000) %>% set_engine("ranger", seed = rngseed) %>% set_mode("regression")
rf_fit1 <- rf_mod1 %>% fit(Y ~ DOSE * AGE * SEX * WT * HT * RACE, data = df1)

lm_fit1
ls_fit1
rf_fit1

lm_pred1 <- predict(lm_fit1, df1) %>% bind_cols(df1)
lm_met1 <- metrics(lm_pred1, truth = Y, estimate = .pred)
lm_met1
ls_pred1 <- predict(ls_fit1, df1) %>% bind_cols(df1)
ls_met1 <- metrics(ls_pred1, truth = Y, estimate = .pred)
ls_met1
rf_pred1 <- predict(rf_fit1, df1) %>% bind_cols(df1)
rf_met1 <- metrics(rf_pred1, truth = Y, estimate = .pred)
rf_met1

lm_p1 <- lm_pred1 %>% ggplot(aes(x=Y, y=.pred)) + geom_point() +ggtitle("Linear model observed vs. predicted")
lm_p1
ls_p1 <- ls_pred1 %>% ggplot(aes(x=Y, y=.pred)) + geom_point() +ggtitle("LASSO model observed vs. predicted")
ls_p1
rf_p1 <- rf_pred1 %>% ggplot(aes(x=Y, y=.pred)) + geom_point() +ggtitle("Random Forest model observed vs. predicted")
rf_p1
```

The RMSE are as follows:
Linear: 475
LASSO: 517
Random forest: 373