---
title: "Fitting Exercise"
---

First to install packages
```{r}
library(here)
library(dplyr)
library(ggplot2)
```

Then to load data
```{r}
data_location <- here::here("fitting-exercise", "Mavoglurant_A2121_nmpk.csv")
df <- read.csv(data_location)
```

Plotting outcome (DV) as a function of time colored by dose and using ID as a grouping factor
```{r}
p1 <- ggplot(df, aes(x=TIME, y=DV, group=ID, color=DOSE)) + geom_line()
plot(p1)
```

remove all OCC=2 entries, keep only OCC=1
```{r}
df2 <- subset(df, (OCC == 1))
```

Exclude all Time=0, found sum of DV for each ID, create a new data frame with only time 0 and a new variable Y that contains amount of drug delivered over the whole experiment
```{r}
df3 <- subset(df2, (TIME != 0))
df4 <- df3 %>% group_by(ID) %>% summarise(Y = sum(DV))
df5 <- subset(df2, (TIME == 0))
df6 <- left_join(x=df5, y=df4, by=c("ID"))
```

Keep only variables Y, Dose, Age, Sex, Race, WT, HT. Convert Race and Sex to factors
```{r}
df7 <- select(df6,Y, DOSE, AGE, SEX, RACE, WT, HT)
df7$SEX <- as.factor(df7$SEX)
df7$RACE <- as.factor(df7$RACE)
```

Plots depicting:
dose and Y colored by Age
Y and weight colored by Sex
boxplot of Y by sex
boxplot of Y by race
```{r}
summary(df7)

p2 <- ggplot(df7, aes(x=DOSE, y=Y, color=AGE)) + geom_point()
plot(p2)

p3 <- ggplot(df7, aes(x=WT, y=Y, color=SEX)) + geom_point()
plot(p3)

p4 <- ggplot(df7, aes(x=SEX, y=Y)) + geom_boxplot()
plot(p4)

p5 <- ggplot(df7, aes(x=RACE, y=Y)) + geom_boxplot()
plot(p5)

hisDOSE <- ggplot(df7, aes(x=DOSE)) + geom_histogram()
plot(hisDOSE)

hisAGE <- ggplot(df7, aes(x=AGE)) + geom_histogram()
plot(hisAGE)

hisWT <- ggplot(df7, aes(x=WT)) + geom_histogram()
plot(hisWT)

hisHT <- ggplot(df7, aes(x=HT)) + geom_histogram()
plot(hisHT)
```

