---
title: "Fitting Exercise"
---

First to install packages
```{r}
library(here)
library(dplyr)
library(ggplot2)
library(tidymodels)
```

Then to load data
```{r}
data_location <- here::here("fitting-exercise", "Mavoglurant_A2121_nmpk.csv")
df <- read.csv(data_location)
```

Plotting outcome (DV) as a function of time colored by dose and using ID as a grouping factor
```{r}
p1 <- ggplot(df, aes(x=TIME, y=DV, group=ID, color=DOSE)) + geom_line()
plot(p1)
```

remove all OCC=2 entries, keep only OCC=1
```{r}
df2 <- subset(df, (OCC == 1))
```

Exclude all Time=0, found sum of DV for each ID, create a new data frame with only time 0 and a new variable Y that contains amount of drug delivered over the whole experiment
```{r}
df3 <- subset(df2, (TIME != 0))
df4 <- df3 %>% group_by(ID) %>% summarise(Y = sum(DV))
df5 <- subset(df2, (TIME == 0))
df6 <- left_join(x=df5, y=df4, by=c("ID"))
```

Keep only variables Y, Dose, Age, Sex, Race, WT, HT. Convert Race and Sex to factors
```{r}
df7 <- select(df6,Y, DOSE, AGE, SEX, RACE, WT, HT)
df7$SEX <- as.factor(df7$SEX)
df7$RACE <- as.factor(df7$RACE)
```

Plots depicting:
dose and Y colored by Age
Y and weight colored by Sex
boxplot of Y by sex
boxplot of Y by race
```{r}
summary(df7)

p2 <- ggplot(df7, aes(x=DOSE, y=Y, color=AGE)) + geom_point()
plot(p2)

p3 <- ggplot(df7, aes(x=WT, y=Y, color=SEX)) + geom_point()
plot(p3)

p4 <- ggplot(df7, aes(x=SEX, y=Y)) + geom_boxplot()
plot(p4)

p5 <- ggplot(df7, aes(x=RACE, y=Y)) + geom_boxplot()
plot(p5)

hisDOSE <- ggplot(df7, aes(x=DOSE)) + geom_histogram()
plot(hisDOSE)

hisAGE <- ggplot(df7, aes(x=AGE)) + geom_histogram()
plot(hisAGE)

hisWT <- ggplot(df7, aes(x=WT)) + geom_histogram()
plot(hisWT)

hisHT <- ggplot(df7, aes(x=HT)) + geom_histogram()
plot(hisHT)
```

fit a linear model to Y using Dose and calculate RMSE/R2
```{r}
lm_mod1 <- linear_reg()
lm_fit1 <- lm_mod1 %>% fit(Y ~ DOSE, data = df7)
lm_fit1
pred1 <- predict(lm_fit1, df7) %>% bind_cols(df7)
met1 <- metrics(pred1, truth = Y, estimate = .pred)
met1
```
In this model dose has a coefficient of 58 which seems like a high number to me. (whenever my collaborators run models we get numbers like 1.2, although those are probably somewhat different models.) The R2 is .516 which is also decent in my opinion. I still have some issues figuring out how to interpret the RMSE and MAE though.


fit a linear model to Y using all predictors and calculate RMSE/R2
```{r}
lm_mod2 <- linear_reg()
lm_fit2 <- lm_mod2 %>% fit(Y ~ DOSE * AGE * SEX * WT * HT * RACE, data = df7)
lm_fit2
pred2 <- predict(lm_fit2, df7) %>% bind_cols(df7)
met2 <- metrics(pred2, truth = Y, estimate = .pred)
met2
```
This one has a lower RMSE/MAE and a higher R2, so I can conclude that it is probably a better model than the first.


fit a linear model to SEX using Dose, and one using all predictors, and calculate ROC-AUC
```{r}
lm_mod3 <- logistic_reg()
lm_fit3 <- lm_mod3 %>% fit(SEX ~ DOSE, data = df7)
lm_fit3
pred3 <- predict(lm_fit3, df7, type = "prob") %>% bind_cols(df7)
roc_auc(pred3, truth = SEX, .pred_1)

lm_mod4 <- logistic_reg()
lm_fit4 <- lm_mod4 %>% fit(SEX ~ DOSE * Y * AGE * WT * HT * RACE, data = df7)
lm_fit4
pred4 <- predict(lm_fit4, df7, type = "prob") %>% bind_cols(df7)
roc_auc(pred4, truth = SEX, .pred_1)
```

The roc_auc is 1 for the last model which is super interesting. I guess that means that all the data put together are very good at predicting sex. This makes sense given that a lot of the things in the model are correlated with sex (height/weight, and doeses tend to be different for men/women). In model 3, dose alone had an ROC of .6, so it's probably a decent predictor of sex.

random seed:
```{r}
rngseed = 1234
```

remove race as a variable
```{r}
df8 <- df7 %>% select(-RACE)
```

set seed and split data 75:25
```{r}
set.seed(rngseed)

data_split  <- initial_split(df8, prop = 3/4)

train <- training(data_split)
test <- testing(data_split)
```

two linear models for Y (continuous outcome of interest)
Model 1: dose as only pred
Model 2: all predictors
```{r}
set.seed(rngseed)
lm_mod5 <- linear_reg()
lm_fit5 <- lm_mod5 %>% fit(Y ~ DOSE, data = train)
lm_fit5
pred5 <- predict(lm_fit5, train) %>% bind_cols(train)
met5 <- metrics(pred5, truth = Y, estimate = .pred)
met5

lm_mod6 <- linear_reg()
lm_fit6 <- lm_mod6 %>% fit(Y ~ DOSE * AGE * SEX * WT * HT, data = train)
lm_fit6
pred6 <- predict(lm_fit6, train) %>% bind_cols(train)
met6 <- metrics(pred6, truth = Y, estimate = .pred)
met6

NM <- null_model(mode = "regression", engine = "parsnip")
NM_fit <- NM %>% fit(Y ~., data = train)
NM_fit
pred7 <- predict(NM_fit, train) %>% bind_cols(train)
met7 <- rmse(pred7, truth = Y, estimate = .pred)
met7
```

I set the random seed correctly, however my values are different from the ones we're supposed to get. I got 948 for the null which is correct, 703 for model 1 (correct value 702), and 578 for model 2 (correct value 627).
This does mean that model 2 is currently performing best as it has the lowest rmse

Now I need to do a 10-fold CV
```{r}
folds <- vfold_cv(train, v = 10)
folds

lm_wf1 <- workflow() %>% add_model(lm_mod5) %>% add_formula(Y ~ DOSE)
lm_fit_cv1 <- lm_wf1 %>% fit_resamples(folds)
lm_fit_cv1
collect_metrics(lm_fit_cv1)

lm_wf2 <- workflow() %>% add_model(lm_mod5) %>% add_formula(Y ~ DOSE * AGE * SEX * WT * HT)
lm_fit_cv2 <- lm_wf2 %>% fit_resamples(folds)
lm_fit_cv2
collect_metrics(lm_fit_cv2)

NM_wf <- workflow() %>% add_model(NM) %>% add_formula(Y ~ .)
NM_fit_cv <- NM_wf %>% fit_resamples(folds)
NM_fit_cv
collect_metrics(NM_fit_cv)
```

Well the rmse for the null model was similar but different at 943, SE at 66.2. RMSE for the second model with the full formula was much worse at 2003, while rmse for the Y~DOSE model was 687.

Now I need a different random seed
```{r}
set.seed(347)
folds2 <- vfold_cv(train, v = 10)
folds2

lm_wf3 <- workflow() %>% add_model(lm_mod5) %>% add_formula(Y ~ DOSE)
lm_fit_cv3 <- lm_wf3 %>% fit_resamples(folds2)
lm_fit_cv3
collect_metrics(lm_fit_cv3)

lm_wf4 <- workflow() %>% add_model(lm_mod5) %>% add_formula(Y ~ DOSE * AGE * SEX * WT * HT)
lm_fit_cv4 <- lm_wf4 %>% fit_resamples(folds2)
lm_fit_cv4
collect_metrics(lm_fit_cv2)

NM_wf2 <- workflow() %>% add_model(NM) %>% add_formula(Y ~ .)
NM_fit_cv2 <- NM_wf2 %>% fit_resamples(folds2)
NM_fit_cv2
collect_metrics(NM_fit_cv2)
```

Null model is now at 938, similar but slightly different
second model is still at 2003
First model is now at 684.
Similar, but slightly different to the other random seed.

# This section added by Rebecca Basta

# Null Model Predictions
```{r}
plot_null <- predict(NM_fit, train) %>%
  bind_cols(train %>% select(Y)) %>%
  mutate(model = "Null model")
```

# Dose Model Predictions
```{r}
plot_mod1 <- predict(lm_fit5, train) %>%
  bind_cols(train %>% select(Y)) %>%
  mutate(model = "Dose only")
```

# All Predictor Model Predictions
```{r}
plot_mod2 <- predict(lm_fit6, train) %>%
  bind_cols(train %>% select(Y)) %>%
  mutate(model = "All predictors")
```

# Combine into One Dataframe
```{r}
plot_df <- bind_rows(plot_null, plot_mod1, plot_mod2)
```

# Predicted vs Observed Plot
```{r}
p_pred_obs <- ggplot(plot_df, aes(x = Y, y = .pred, color = model, shape = model)) +
  geom_point(alpha = 0.7, size = 2) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  coord_cartesian(xlim = c(0, 5000), ylim = c(0, 5000)) +
  labs(
    x = "Observed values",
    y = "Predicted values",
    title = "Predicted vs observed values for 3 training-set models"
  ) +
  theme_minimal()

plot(p_pred_obs)
```


The model that uses dosee as the only predictor, the predicted values fall along three horizontal lines becuase thedose cariable takes three values. 

#  Plot with Model Using All Predictors 
```{r}
resid_df <- predict(lm_fit6, train) %>%
  bind_cols(train %>% select(Y)) %>%
  mutate(residual = .pred - Y)

p_resid <- ggplot(resid_df, aes(x = .pred, y = residual)) +
  geom_point(alpha = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  coord_cartesian(ylim = c(-2000, 2000)) +
  labs(
    x = "Predicted values",
    y = "Residuals (predicted - observed)",
    title = "Predicted values vs residuals for model 2"
  ) +
  theme_minimal()

plot(p_resid)
```

# Model Predictions and Uncertainty 
```{r}
library(rsample)
set.seed(rngseed)

# Create 100 bootstrap resamples
dat_bs <- bootstraps(train, times = 100)
```

```{r}
# Store predictions
n <- nrow(train)
B <- 100
pred_bs <- matrix(NA, nrow = B, ncol = n)

# Loop through bootstrap samples
for (i in 1:B) {
  
  dat_sample <- rsample::analysis(dat_bs$splits[[i]])
  
# fit model on bootstrap sample
  fit_bs <- lm(Y ~ DOSE * AGE * SEX * WT * HT, data = dat_sample)
  
# predict on original training data
  pred_bs[i, ] <- predict(fit_bs, newdata = train)
}
```

# 4. Compute Median and 89% Confidence Intervals 
```{r}
preds <- pred_bs |> apply(2, quantile, c(0.055, 0.5, 0.945)) |> t()

# Original predictions
orig_pred <- predict(lm_fit6, train)

# Put in dataframe
boot_df <- data.frame(
  observed = train$Y,
  pred_mean = orig_pred,
  pred_lower = preds[,1],
  pred_median = preds[,2],
  pred_upper = preds[,3]
)
```

# Plot Predicted vs Observed with Uncertainty
```{r}
ggplot() +
  geom_point(data = boot_df, aes(x = observed, y = .pred), color = "black", alpha = 0.7) +
  geom_point(data = boot_df, aes(x = observed, y = pred_median), color = "blue", alpha = 0.7) +
  geom_errorbar(
    data = boot_df,
    aes(x = observed, ymin = pred_lower, ymax = pred_upper),
    color = "red",
    alpha = 0.5
  ) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  coord_cartesian(xlim = c(0, 5000), ylim = c(0, 5000)) +
  labs(
    x = "Observed values",
    y = "Predicted values",
    title = "Bootstrap predictions with 89% confidence intervals"
  ) +
  theme_minimal()
```

The black points represent the original model predictions, and most of the points lie close to the 45° line, this indicates that there is not much variability between observed and predicted values. The blue points represent the median predictions from the bootstrap samples, they overlap with the original predictions. This overlap shows that the model is consistent across resampled datasets. The red vertical lines indicate the 89% confidence intervals, which are small for most individuals demonstrating low uncertainty, but are bigger for a few individuals. The pattern shows that the model represents most of trends in the data, and the bootstrap confirms the consistency and uncertainty of the predictions.