---
title: "Tidy Tuesday"
---
This cell loads packages and the required files
```{r}
library(dplyr)
library(ggplot2)
library(tidymodels)
library(readr)
library(skimr)
library(forcats) #for collapsing factors
beaufort_scale <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-04-14/beaufort_scale.csv')
birds <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-04-14/birds.csv')
sea_states <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-04-14/sea_states.csv')
ships <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-04-14/ships.csv')
```

Now I want to look at these files, it seems like there's four of them

```{r}
skim(beaufort_scale)

skim(birds)

skim(sea_states)

skim(ships)
```

Ok so there's a lot here. I need to find a reasonable question to ask. I am trying to figure out if there is any overlap in variables between the datasets. For example it would be interesting to see if like one bird was seen more often on windy days or something.
I think the record_id variable is an overlap between the birds and the ships
according to the github it is and also: record_id 1184009, bird_observation_id 974 has no corresponding ship id
Meanwhile the other two datasets are just scales that can maybe help interpret the data.

First to clean the data: birds and ships have several variables that have so much missing data that they're useless so I want to remove them
```{r}
birds2 <- birds %>% dplyr::select(-wan_plumage_phase, -plumage_phase, -sex, - moulting, -age)
ships2 <- ships %>% dplyr::select(-direction, -wind_direction, -air_temperature, -pressure, -sea_surface_temperature, -depth)

skim(birds2)
skim(ships2)
```

I like birds. I want to know under which weather conditions I am most likely to get birds accompanying my ship.
I will use n_accompanying as my outcome variable and weather conditions measured by the ships as the predictors.

First I need to do more cleaning. There are a lot of NA's in this data. I am going to use the logical variable, accompanying to help determine what to keep. If accompanying = FALSE and n_accompanying = NA then I will set n_accompanying to 0. If accompanying = NA, or accompanying = TRUE and n_accompanying = NA, then I will remove the datapoint. I can't tell if under the first circumstance there were no accompanying, or if they forgot to check. Meanwhile under the second I can't tell how many, so I will remove it.

```{r}
whole <- merge(birds2, ships2, by="record_id")
#merge the data all together!

#select my dataset
df <- whole %>% select(bird_observation_id, record_id, accompanying, n_accompanying, speed, cloud_cover, precipitation, wind_speed_class, sea_state_class, season)

#get rid of anything where it is unclear if accompanying ship
df <- df %>% drop_na(accompanying)

#drop rows where accompanying = true, n_accompanying = NA
df2 <- df %>% dplyr::filter(accompanying == TRUE & is.na(n_accompanying))
df3 <- dplyr::anti_join(df, df2, by = "bird_observation_id")

#make 0 values where accompanying = False n_accompanying = NA
df3$n_accompanying <- replace_na(df3$n_accompanying, 0)

#I also need to remove any where accompanying == TRUE and n_accompanying = 0
df4 <- df3 %>% dplyr::filter(accompanying == TRUE & n_accompanying == 0)
df5 <- dplyr::anti_join(df3, df4, by = "bird_observation_id")

#There are a lot of 0s and I want to be able to look at things without being overwhelmed by them so I will make another df excluding them
df6 <- df5 %>% filter(accompanying == TRUE)
df7 <- df5 %>% filter(accompanying == FALSE)
summary(df6)
summary(df7)
#This shows that accompanying now matches n_accompanying correctly
```

Now I want to do some EDA for n_accomanying
```{r}
p1 <- ggplot(df6, aes(x=n_accompanying)) + geom_histogram()
p1
#This looks like I have some outliers in my data, they could be real, esp if you have really large flocks of birds. But it does make things a little difficult to see. For simplicities sake I will remove everything over 250 birds. No idea if this is correct or not

df8 <- df5 %>% filter(n_accompanying < 250)
summary(df8)
#it looks like some things that should be categorical are strings

df8$season <- as.factor(df8$season)
df8$cloud_cover <- as.factor(df8$cloud_cover)
df8$precipitation <- as.factor(df8$precipitation)

#it seems like for precipitation, we can get rid of some of the NA's by looking at cloud cover. 
#If cloud cover = clear we can assume there is none precipitaion
df9 <- df8 %>% dplyr::filter(cloud_cover == "clear" & is.na(precipitation))
df9$precipitation <- replace_na(df9$precipitation, "none")
df10 <- rows_update(df8, df9, by = "bird_observation_id")

#I want to combine some of the precipitation levels for simplicities sake
#Specifically into: none, rain, snow, other

df10$precipitation <- fct_collapse(df10$precipitation, rain = c("rain", "showers", "squalls", "drizzle"), snow = c("snow showers", "continuous snow"))
summary(df10)
#data looks pretty good I think!

df6 <- df10 %>% filter(accompanying == TRUE)
```

Actual EDA time!
```{r}
p2 <- df6 %>% ggplot(aes(x=speed, y=n_accompanying)) + geom_point()
p2

p3 <- df6 %>% ggplot(aes(x=wind_speed_class, y=n_accompanying)) + geom_point()
p3

p4 <- df6 %>% ggplot(aes(x=sea_state_class, y=n_accompanying)) + geom_point()
p4

p5 <- df6 %>% ggplot(aes(x=cloud_cover, y=n_accompanying)) + geom_boxplot()
p5

p6 <- df6 %>% ggplot(aes(x=precipitation, y=n_accompanying)) + geom_boxplot()
p6
```

Looks like theres more birds maybe when there's less precipitation
Also sea state might be a thing

Now time to get modeling
My three model types will be:
Generalized linear model
Random Forest model
?

To start I will split the data and set a seed
```{r}

```

Here is my linear model
```{r}

```

Here is my random forest model
```{r}

```

Here is my X model
```{r}

```

I plan to choose the XXXX as my model because.....

Here I will use my test data and evaluate it
```{r}

```